Bayesian methods for black hole merger parameter estimation with LISA.

### 2019-04-04 paper/paper_PE.tex
Updates to SOBH section

-- Fisher matrices computation: step self-tuning,...

-- Bayesian samplers: Multinest, PTMCMC

@@

\section{Stellar origin black holes}
\label{sec:SOBH}

@@

Stellar-origin black-hole (SOBH) mergers have been recently recognized as an
important source of LISA detections \cite{Sesana16} although full parameter
estimation studies are not available yet. In this section we present
three case studies of parameter inference for such sources. The parameters of
the simulated mergers are taken from the Radler LISA data challenge and are
listed in Table \ref{table:sobh_params}.
\begin{table}
	\begin{tabular}{|c||c|c|c|}
		\hline
		Radler entry 		& 1 	& 12 	& 52 \\
		\hline
		Mass 1 ($\Msol$) 	& 21.4	& 63.5	& 57.0 \\
		\hline
		Mass 2 ($\Msol$) 	& 20.1	& 45.1	& 36.8 \\
		\hline
		Distance (Mpc) 		& 49.1	& 318	& 308 \\
		\hline
		LISA SNR 			& 26.6	& 11.9	& 16.0 \\
		\hline
	\end{tabular}
	\caption{Parameters of the simulated SOBH mergers. \tdc{Maybe we can show
			 these on the upper right of each corner plot instead.}}
	\label{table:sobh_params}
\end{table}

The waveform model used to work with such systems is the phenomenological
approximant IMRPhenomD, extended at low frequency with a post-Newtonian
frequency-domain waveform. In practice, because LISA only observes the early
inspiral of SOBH systems, only the post-Newtonian waveform plays a role here.
We do not expect radiation modes beyond the quadrupole to give a significant
contribution to the parameter estimates, so we only include the $(2,2)$ mode
for simplicity. We do not include any effect arising from black-hole spins.

@@

>>>>

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Stellar origin black holes}
\label{sec:SOBH}

@@

Stellar-origin black-hole (SOBH) mergers have been recently recognized as an
important source of LISA detections \cite{Sesana16} although full parameter
estimation studies are not available yet. In this section we present two case
studies of parameter inference for such sources using the formalism and methods
described earlier. The parameters of the simulated mergers are taken from the
Radler LISA data challenge and are listed in Table \ref{table:sobh_params}.
\begin{table}
	\begin{tabular}{|c||c|c|}
		\hline
		Radler entry 		& 1 	& 12   \\
		\hline
		Mass 1 ($\Msol$) 	& 21.4	& 63.5 \\
		\hline
		Mass 2 ($\Msol$) 	& 20.1	& 45.1 \\
		\hline
		Distance (Mpc) 		& 49.1	& 318  \\
		\hline
		LISA SNR 			& 26.6	& 11.9 \\
		\hline
	\end{tabular}
	\caption{Parameters of the simulated SOBH mergers. \tdc{Maybe we can show
			 these on the upper right of each corner plot instead.}}
	\label{table:sobh_params}
\end{table}

We caution the reader that our purpose here is to show the feasibility of
Bayesian inference for SOBH sources in LISA using the methodology presented in
this paper. It is by no means a final answer on how to do so in an optimal or
convenient way.  For this reason, the problem is posed in a simplified
way here: in particular, the SOBH signal is assumed to merge roughly at the end of
the observation epoch, and the finite observation time is enforced by
converting it to a minimum analyzed Fourier frequency via the time-frequency
relation of a simple Newtonian inspiral. In a more realistic situation, the
merger time would likely be free to range from the beginning of the observation
to years after its end, and the frequency of the signal at the beginning of the
observation would be another sampling parameter.

The waveform model used to work with SOBH systems is the same phenomenological
approximant PhenomD used earlier, extended at low frequency with a post-Newtonian
frequency-domain waveform. In practice, because LISA only observes the early
inspiral of SOBH systems, only the post-Newtonian waveform plays a role here.
We do not expect radiation modes beyond the quadrupole to give a significant
contribution to the parameter estimates, so we only include the $(2,2)$ mode
for simplicity. We do not include any effect arising from black-hole spins.

@@

Such posterior distributions should not be taken literally as expectations of
LISA's capability for SOBH systems. One reason is the simplified treatment of
the finite observation time described earlier. A second reason is the assumption
of zero black-hole spins, which artificially reduces the width of the posterior
distribution due to the strong degeneracy between spins and symmetric mass
ratio \cite{Baird+2013}.
